---
title: "ST 308 Project"
author: "Hung Nguyen, Landrey Lord"
format: pdf
---

## Part 1
```{r}
library(tidyverse)
```
### Question 1
Quarto is a type of document in R that allows us to make documents or reports 
that can contain text, code, and show output all in one. Quarto document is 
useful for people who work with data because it saves the codes written, 
allowing people to go back and make any changes if needed. Quarto also helps us render to pdf file containing the graphical outputs of our codes, which is 
useful for making reports (This question is answered by Landrey Lord)
```{r}
```
### Question 2
```{r}
library(readr)
house <- read_csv("aacline_house.csv")
head(house)
```
```{r}
```
(This question is answered by Hung Nguyen)
```{r}
```
### Question 3
GarageCars is a quantitative column, which has numerical values and specifically
integer values in our dataset. HouseStyle is a categorical column, which has character values in our dataset (This question is answered by Hung Nguyen)
```{r}
```
### Question 4
```{r}
house %>%
  filter(YearBuilt > median(YearBuilt),
         SalePrice < mean(SalePrice),
         HouseStyle == "2Story")
```
(This question is answered by Landrey Lord)
```{r}
```
### Question 5
```{r}
house_subset <- 
  house %>%
    rename(sale_price = SalePrice,
           garage_cars = GarageCars,
           year_built = YearBuilt) %>%
    mutate(garage_cars = factor(garage_cars)) %>%
    select(sale_price, garage_cars, year_built)
head(house_subset)

```
(This question is answered by Hung Nguyen)
```{r}
```
### Question 6
```{r}
tapply(X = house$WoodDeckSF, INDEX = list(house$Foundation, house$RoofStyle), 
       FUN = mean, na.rm = TRUE)
tapply(X = house$WoodDeckSF, INDEX = list(house$Foundation, house$RoofStyle), 
       FUN = sd, na.rm = TRUE)
```
(This question is answered by Hung Nguyen and Landrey Lord)
```{r}
```
## Part 2
```{r}
```
### Question 1
In the dataset, each observation has 7 attributes. Most variables are 
categorical, including START_DATE, END_DATE, CATEGORY, START, STOP, PURPOSE. 
Only MILES is a quantitative variable. In addition, START_DATE and END_DATE 
variables range from January to December 2016. Meanwhile, CATEGORY has two 
values: ‘Business’ and ‘Personal’. START and STOP variables have values of 
places in the United States. MILES are quantitative values with many different travelled distances. PURPOSE contains specific trip reasons though many 
observations have missing values for this variable. The data came from a 
driver's personal Uber ride history. Every ride recorded by Uber includes: start
time, end time, pickup location, drop-off location, miles traveled, trip purpose
(if added manually). The research question is which are the top 3 most common travelling purposes, whether it is personal or business category, 
and which one has the longest average travelled distance. This research question relates to three variables: CATEGORY, PURPOSE, MILES. It answers which of the travelling purposes is popular and determines whether it is either a personal or business category. Audiences also can see which travelling purposes motivates 
people to travel longer distances by Uber, which costs them a considerable 
amount of money (This question is answered by Hung Nguyen and Landrey Lord)
```{r}
```
### Question 2
```{r}
uber_data <- read_csv("UberDataset.csv")

uber <- uber_data %>%
  mutate(PURPOSE = na_if(PURPOSE, "")) %>% 
# na_if() function is used to convert blank values in PURPOSE column into NA
# values to serve for drop_na() function and then perform statistical summary
  drop_na(PURPOSE)

table(uber$PURPOSE, uber$CATEGORY)

uber %>%
  group_by(PURPOSE, CATEGORY) %>%
  summarise(miles_avg = mean(MILES))

library(ggplot2)
ggplot(uber, aes(x = PURPOSE, y = MILES)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
# we used this theme() function to format the x-axis of our boxplot to make it
# easier to see values on the x-axis
  labs(
    title = "Boxplot for distribution of miles travelled by travelling 
    purposes",
    x = "Travelling Purposes",
    y = "Miles travelled"
  )
```
The most common travelling purpose is Meeting, the second one is Meal/Entertain, 
the third one is Errand/Supplies, all of these three purposes fall under the 
Business category. However, the highest average miles travelled is Commute which falls under the Personal category. The data presents that the top three most 
common ways of travelling are Meeting, Meal/Entertain, and Errand/Supplies, 
which belong to the Business category, none of them have the highest average 
miles travelled like we thought. The main reason behind this is that people do 
not travel long distances for these purposes, so the average miles travelled is 
not considerably high. With that people travel longer distances for Commute 
purposes, so the miles travelled is significantly high compared to the low 
frequency. With this specific case there is only 1 person travelling for 
Commute and that person travelled over 180 miles, making this the highest number 
we recorded of average miles travelled (This question is answered by Hung Nguyen 
and Landrey Lord)









